---
title: "Scrape GEPRIS"
author: "Till Grallert"
format: html
editor: visual
---

Initially it seemed simple to scrape all entries from the database as the URLs follow the pattern `https://gepris.dfg.de/gepris/projekt/{ID}` with numerical IDs. However, as examples showed that these can have at least 9 digits, a simple incrementing function would take more than 25 years with a one second delay between each call to the server in order to scrape the 135273 actual projects listed in GEPRIS (because IDs do not form an uninterrupted sequence of numbers this attempt would make many unnecessary calls).

The first step, therefore, was to scrape the existent project IDs from the catalogue pages. In total this would require to iterate over 2706 catalogue pages with 50 projects IDs each but could be further decreased by limiting our call to a specific sub-field, namely the humanities, encoded in the catalogue URLs as `&fachgebiet=11`. This would reduce the number of catalogue pages to 319.

# prepare the environment

In order to scrape websites and extract information based on their position within the parsed HTML5 / XML, we need to load the `rvest()` library.

```{r}
library(tidyverse)
library(rvest)
library(here)
```

We also need a small helper function to extract the textual content of any element or attribute within a given HTML stream.

```{r}
# input for the xpath parameter is a string
f.get.text <- function(html, xpath) {
  v.text = rvest::html_elements(html, xpath = xpath) %>%  
    rvest::html_text2()
  # test if output is of zero-length
  #print(v.text)
  ifelse( length(v.text) == 0, 
          "",
          v.text)
}
```

# Scrape the relevant project IDs from the catalogue

The following function scrapes all project IDs from a GEPRIS catalogue page.

```{r}
f.scrape.gepris.ids <- function(url) {
  v.page = rvest::read_html(url)
  v.xpath.item = "/html/body//div[@class = 'listbox2']/div[@id = 'liste']/div"
  v.urls.projects <- ''
  for (item in html_elements(v.page, xpath = v.xpath.item)) {
    v.urls.projects <- append(v.urls.projects,
                            f.get.text(item, "div[@class = 'results']/h2[1]/a/@href"),
                            after = 1)
  }
  str_replace(v.urls.projects, '^.*?(\\d+)$', '\\1')
}
```

We can now use this function iterate over the catalogue pages using our knowledge of the URL patterns:

-   '&index=0' needs to be incremented by 50
-   '&fachgebiet=11' indicates humanities

```{r}
## parts of the URL as variables. 
v.url.base = "https://gepris.dfg.de/gepris/OCTOPUS"
v.url.params = "findButton=historyCall&context=projekt&einrichtungsart=-1&nurProjekteMitAB=false&pemu=%23&teilprojekte=true&zk_transferprojekt=false&hitsPerPage=50"
v.url.param.area = "fachgebiet=" # 11: humanities
v.url.param.index = "index="
v.url.param.task = "task=doKatalog"        # doSearchExtended, doKatalog
v.url.param.country = "continentId=5&countryKey=ISR" # ISR: Israel. I haven't got this to work yet
```

We also ought to be kind to the server and add some delay (unfortunately their really rudimentary robots.txt did not provide any guidance for crawlers). After that we can crawl all the project IDs, we are interested in, and save them to file.

NOTE: Do NOT execute more than once, as it will potentially try and crawl tens of thousands of IDs.

```{r scrape-gepris-ids, eval=FALSE, include=FALSE}
# set delay in seconds
v.delay = 0.1

# iterate over catalogue pages with 50 results each to scrape all project IDs
v.ids.project = c('')
v.onset = 0
v.terminus = 1000 #319 # terminus depends on the number of expected results.
for (page in v.onset:v.terminus) {
  print(paste('scrape GEPRIS IDs: page', page + 1, 'of', v.terminus + 1))       # progress message
  v.url = paste0(v.url.base, '?', v.url.params, '&', v.url.param.task,           # compose target URL
                 #'&', v.url.param.area, '11',                                   # limit to a field: humanities
                 '&', v.url.param.index, page * 50) 
  v.ids.project <- append(v.ids.project, f.scrape.gepris.ids(v.url), after = 1) # scrape URL and amend list of IDs
  Sys.sleep(v.delay)                                                            # delay
}

# convert to tibble
df.ids.project <- tibble(id = v.ids.project) %>%
  dplyr::mutate(id = na_if(id, '')) %>%
  tidyr::drop_na() %>%
  dplyr::arrange(id)

# save to disk
write.table(df.ids.project, file = here("data/dfg/dfg_project-ids.csv"), row.names = F, sep = ",")
```

# get individual project details

Functions to scrape project descriptions, applicants, and collaborators from individual project pages on GEPRIS

```{r}
f.scrape.gepris.project.desc <- function(project.id, delay) {
  v.url.base = "https://gepris.dfg.de/gepris/projekt/"
  v.url = paste0(v.url.base, project.id, '?context=projekt&task=showDetail&id=', project.id)
  #print(v.url)
  v.page = rvest::read_html(v.url)
  Sys.sleep(delay) # add delay as courtesy to the hosting servers and allow time to fully load the page
  v.xpath.details = "/html/body//div[@class = 'details']"
  v.xpath.description = "/html/body//div[@id = 'projektbeschreibung']"
  #print(f.get.text(v.page, v.xpath.details))
  df.details <- dplyr::tibble(
    project.id = as.character(project.id),
    # data from "details"
    project.title = f.get.text(v.page, paste(v.xpath.details, "h1[1]", sep = "/")),
    discipline = f.get.text(v.page, paste(v.xpath.details, "span[@class = 'name'][text() = 'Fachliche Zuordnung']/following-sibling::span[@class = 'value']", sep = "//")),
    dates = f.get.text(v.page, paste(v.xpath.details, "span[@class = 'name'][text() = 'Förderung']/following-sibling::span[@class = 'value']", sep = "//")),
    # data from description
    project.desc = f.get.text(v.page, paste(v.xpath.description, "div[@id = 'projekttext']", sep = "/")),
    programme = f.get.text(v.page, paste(v.xpath.description, "div[@id = 'projekttext']/following-sibling::div[1]/span[@class = 'value']", sep = "/"))
    # international cooperation: country
    #collaborator.country = f.get.text(v.page, paste(v.xpath.description, "div[@id = 'projekttext']/following-sibling::div[2]/span[@class = 'value']", sep = "/"))
  ) %>%
    # process dates
  mutate(onset = as.double(str_replace(dates, '^.*?(\\d{4}).*$', '\\1')),
         terminus = as.double(str_replace(dates, '^.*(\\d{4})*.*(\\d{4}).*?$', '\\2')))
  
  df.details
}
# create an empty placeholder dataframe 
df.projects <- data.frame(
  project.id = character(),
  project.title = character(),
  discipline = character(),
  dates = character(),
  onset = double(),
  terminus = double(),
  project.desc = character(),
  programme = character(),
  collaborator.country = character()
)
```

Test the scraping functions with a complex project

```{r test-scrape-gepris-project}
v.project = '158416540' #'100000777' # '158416540'
f.scrape.gepris.project.desc(v.project, 1)
```

```{r}
f.scrape.gepris.project.applicants <- function(project.id) {
  v.url.base = "https://gepris.dfg.de/gepris/projekt/"
  v.url = paste0(v.url.base, project.id)
  v.page = rvest::read_html(v.url)
  v.xpath.details = "/html/body//div[@class = 'details']"
  dplyr::tibble(
    project.id = as.character(project.id),
    # data from "details"
    applicant.id = str_replace(
      f.get.text(v.page, paste(v.xpath.details, "span[@class = 'value']/a[starts-with(@href, '/gepris/person')]/@href", sep = "//")),
      '^.*?(\\d+)$', '\\1'),
    applicant.name = f.get.text(v.page, paste(v.xpath.details, "span[@class = 'value'][starts-with(a/@href, '/gepris/person')]/a", sep = "//"))
  )
}
# create an empty placeholder dataframe 
df.applicants <- data.frame(
  project.id = character(),
  applicant.id = character(),
  applicant.name = character()
)
```

Test the scraping functions with a complex project

```{r test-scrape-gepris-project}
v.project = '100000777' # '158416540'
f.scrape.gepris.project.applicants(v.project)
```

get additional details

```{r}
f.scrape.gepris.project.collaborators <- function(project.id) {
  v.url.base = "https://gepris.dfg.de/gepris/projekt/"
  v.url = paste0(v.url.base, project.id)
  v.page = rvest::read_html(v.url)
  v.xpath.description = "/html/body//div[@id = 'projektbeschreibung']"
  dplyr::tibble(
    project.id = as.character(project.id),
    collaborator.id = str_replace(
      f.get.text(v.page, paste(v.xpath.description, 
            "div/span[@class = 'name'][contains(text(), 'Beteiligte Personen')]/following-sibling::span[@class = 'value']/a[starts-with(@href, '/gepris/person')]/@href", sep = "/")),
      '^.*?(\\d+)$', '\\1'),
    collaborator.name = f.get.text(v.page, paste(v.xpath.description, 
            "div/span[@class = 'name'][contains(text(), 'Beteiligte Personen')]/following-sibling::span[@class = 'value']/a[starts-with(@href, '/gepris/person')]", sep = "/"))
  )
}
# create an empty placeholder dataframe 
df.applicants <- data.frame(
  project.id = character(),
  collaborator.id = character(),
  collaborator.name = character()
)
```

Test the scraping functions with a complex project

```{r test-scrape-gepris-project}
f.scrape.gepris.project.collaborators(v.project)
```

# iterate over projects

1.  Iterate over a range of project IDs (or all of them) and scrape the details for each.
2.  Scrape the details for each involved person based on a list of unique IDs from those involved in the scraped projects

## 1. Load project IDs

```{r load-grepris-ids, include=FALSE}
# load DFG project IDs from file
df.ids.project <- read_csv(here("data/dfg/dfg_project-ids.csv")) %>%
  dplyr::rename(project.id  = id) %>%
  dplyr::mutate(project.id = as.character(project.id)) %>%
  dplyr::arrange(project.id) %>%
  unique()
```

## 2. Iterate over project IDs

We have already created empty dataframes to be filled iteratively.

```{r}
# create empty tibbles
v.project = '100000777'
df.projects <- f.scrape.gepris.project.desc(v.project, 1)#%>%
  dplyr::rows_delete(tibble(project.id = v.project))
df.applicants <- f.scrape.gepris.project.applicants(v.project) %>%
  dplyr::rows_delete(tibble(project.id = v.project))
df.collaborators <- f.scrape.gepris.project.collaborators(v.project) %>%
  dplyr::rows_delete(tibble(project.id = v.project))
```

```{r}
v.onset = 1
v.terminus = 2
for(project.id in df.ids.project[v.onset:v.terminus, ]$project.id) {
  print(paste0("try and scrape DFG project no.", project.id, " from GEPRIS"))
  # test if the project ID is already present in the target tibble
  ifelse(project.id %in% df.projects$project.id,
         print(paste0('project no.', project.id, ' has already been scraped')),
    #  else scrape data for project ID
     df.projects <- df.projects %>%
        dplyr::add_row(f.scrape.gepris.project.desc(project.id, 0.5))
  )
  #case_when( project.id %in% df.projects$project.id ~ print(paste0("Scraping DFG project no.", project.id, " from GEPRIS")))
}
head(df.projects, 5)
```

```{r eval=FALSE, include=FALSE}
# iterate through a range of project IDs
df.dfg.projects = f.scrape.gepris.project(df.ids.project[1, "id"])
v.onset = 2
# set the terminus: for testing purposes this could be "v.onset + 10"
v.terminus = nrow(df.ids.project)
for(id in df.ids.project[v.onset:v.terminus, ]$id) {
  print(paste("Scraping DFG project no.", id, "from GEPRIS", sep = " "))
  df.dfg.projects <- df.dfg.projects %>%
    add_row(f.scrape.gepris.entry(id))
  Sys.sleep(0.5)
}

# save output
save(df.dfg.projects, file = "dfg_projects.rda")
write.table(df.dfg.projects, file = "dfg_projects.csv", row.names = F, sep = ",")
```
